Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-08-01 paper.tex
Updates in analysis and implementation

% packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsfonts} % for R symbol (the set of real numbers)
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{stmaryrd} % for llbracket and rrbracket
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

@@

@@

\subsection{Implementation}
\label{sec:implementation}

Our algorithms are implemented in Python using the Nibabel
library. \todo{Val\'erie and Yongping: (1) extract ImageUtils from
  mapreduce repository (preserve git history; just copy the repo and
  remove the mapreduce stuff in the copy). Link to the github
  repo. (2) Describe the code briefly, in particular the
  implementation tricks used for each implementation.}

\begin{itemize}
\item In MR, no seeking withing the blocks.
\item Avoid in-memory copies.
\end{itemize}

\subsubsection{Cluster reads}

\subsubsection{Multiple reads}

For faster reading all the blocks, before the merging process start,
we read all the split’s headers to get their position information and
calculate each row’s index in the reconstructed image. We only read the
split when there is a row in that writing range (available memory),
otherwise, we skip it.

Initially, we use a numpy array as our data buffer, after the split is read,
we assign the rows (a numpy array of unsigned integers) into this
buffer to the right position. However, after some tests, we found that
there is some overhead. Each time we assign the data to the corresponding
position of the buffer, the system will copy these integers first.
This is neither time-efficient nor space-efficient.

We then think of a better solution which we use the dictionary in python
as our data buffer. We use the row’s index as a key, the data (in bytes)
as the value. As for the multiple reads, the data buffer only contains
contiguous bytes, we can just seek once (the first key of the sorted keys
in dictionary), then write all the bytes continuously to the file.
There is no copy in the memory, which proves a bit faster than using
a numpy array as the buffer.

\todo{Yongping: among other implementation techniques (e.g. dictionary
  vs list, etc), explain clearly and concisely what the calculation
  time covers.}

\section{Simulations}

After running a few simulations, we saw that the comparison between
Cluster reads and Multiple reads essentially depends on the three
Cluster reads configurations.  When Cluster reads merge incomplete
block rows (case 1), they are always outperformed by Multiple reads.
When Cluster reads merge complete block rows (case 2), the comparison
between Cluster reads and Multiple reads depends on the values of $D$
and $n$. Figure~\ref{fig:model-comparison} shows two situations where
Cluster reads and Multiple reads are alternatively better than each
other. When Cluster reads merge complete block slices (case 3),
Cluster reads and Multiple reads are equivalent.

\begin{figure}
  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-creads.pdf}
  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-mreads.pdf}
  \caption{Number of seeks for Cluster reads (Left) and Multiple reads (Right), for
    D=3000, b=2 and n=125.}
  % Gnuplot file: scripts/model/model.gnplt
  \label{fig:model-individual}
\end{figure}

\section{Benchmark}
\label{sec:benchmark}

>>>>

% packages
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsfonts} % for R symbol (the set of real numbers)
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=black,filecolor=black,urlcolor=blue}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{stmaryrd} % for llbracket and rrbracket
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

@@

\subsection{Analysis}

Based on a few simulations of Equations~\ref{eq:seeks-cluster-reads-1}
and~\ref{eq:seeks-multiple-reads}, we conjectured that the comparison
between Cluster reads and Multiple reads essentially depends on the
three Cluster reads cases. When Cluster reads merge incomplete block
rows (case 1), they are always outperformed by Multiple reads
(\textbf{prop 1}).  When Cluster reads merge complete block rows (case
2), the comparison between Cluster reads and Multiple reads depends on
the values of $D$ and $n$ (\textbf{prop 2}). When Cluster reads merge complete
block slices (case 3), Cluster reads and Multiple reads are equivalent
(\textbf{prop 3}).

\textbf{prop 1} \todo{do it!}

 \textbf{prop 2} is proven by Figure~\ref{fig:model-comparison} that
 shows two situations where Cluster reads and Multiple reads are
 alternatively better than each other.

 \textbf{prop 3} is easy to infer from
 equations~\ref{eq:seeks-cluster-reads-1}
 and~\ref{eq:seeks-multiple-reads} when assuming that $Rb$ is an exact
 multiple of $m$. In this case, both equations boils down to
 $n+\frac{Rb}{m}$.

\section{Implementation}
\label{sec:implementation}

The 5 algorithms presented earlier are implemented in a Python library
called \texttt{sam} (for ``split and merge''), available at
\url{https://github.com/big-data-lab-team/-bin-sam} under MIT
license. It uses Nibabel~\cite{matthew_brett_2016_60808} for image I/O
and NumPy for array manipulations.

% buffer implementation
The data buffer used in Cluster reads and Multiple reads is
implemented as a Python dictionary where the keys are offsets in the
reconstructed image and the values are NumPy arrays containing the
data starting at this offset. When the memory load is complete,
dictionary entries are written sequentially to the reconstructed
image. In Cluster reads, some seeking might be required between
writes. We implemented a defragmentation procedure for the dictionary
that merges contiguous dictionary entries in a single one, but we
abandoned it as it proved more time-consuming than going through all
the initial entries, due to the overhead of resizing NumPy arrays to
merge entries. In Multiple reads, dictionary entries are always
contiguous in the reconstructed image.  We tried to use a single NumPy
array as a buffer, but we finally abandoned it as inserting data at a
specific position in a NumPy array copies the data in memory, which
increases the runtime and the peak memory consumption.

% Split methods
The implementation of splitting algorithms was greatly facilitated by
the availability of so-called array proxies in Nibabel, which help to
read specific sub-parts of large images. Nibabel's array proxies
essentially provided the buffer implementation for splitting
algorithms. Unfortunately, they are not available to write
data. \todo{Yongping, Valerie: any other comment on split
  implementations?}

% Additional optimizations.
In Multiple reads, for faster reading all the blocks, before the
merging process start, we read all the block headers to get their
position information and calculate each corresponding row index in the
reconstructed image. We only read the block when there is a row in
that writing range (available memory), otherwise, we skip it.
\todo{Valerie, Yongping, any other implementation trick to report?}

@@
